---
title: "Drug assay preprocessing"
output: html_notebook
---

## Paths

**Copy Paste Here the path to your Drug-Assay-Study folder**

```{r}
my_path<-"/Users/jonasamar/Desktop/Drug-Assay-Study"
```

```{r}
setwd(paste0(my_path,"/Rscripts"))
funct_path=paste0(my_path,"/Drug assay data/Drug Assay Project Initial Data/Drug assay ID1-4re-normalized&barcorded - Statistics - ID2P3 removed.csv")
path_to_penalization_file = paste0(my_path, "/Drug assay data/Drug Assay Project Initial Data/penalization matrix drug assay.xlsx")
out_path=paste0(my_path,"/Drug assay data/Preprocessed Data")
```

Creating directory to store the preprocessed data

```{r}
if (!file.exists(out_path)){dir.create(out_path)}
```

## Libraries

```{r}
library(dplyr)
library(tidyr)
library(stringr)
library(readxl)
```

## Restructuring of data

Importing the data.

```{r}
MyData = read.csv(funct_path)
head(MyData)
```

Removing redundant information (channel/reagent, uniquePopulationName/population) and useless columns (filename, parentPopulation).

```{r}
MyData = subset(MyData, select = -c(filename, uniquePopulationName, parentPopulation, channel))
```

Rewriting the Drug used on each Plate.

```{r}
#This vector contains the name of the drugs in the order corresponding to the plate number they were tested
Drugs <- c( 
  "Cefotaxime",
  "Lansoprazole",
  "Iopamidol",
  "Iohexol",
  "Benzylpenicillin",
  "Chlorthalidone",
  "Rifabutin",
  "Iodixanol",
  "Metformin",
  "Folic acid",
  "Clotrimazole",
  "Maprotiline",
  "Progesterone",
  "Pravastatin",
  "Methylpredonisolone"
)

# We associate the name of the drug corresponding to its plate number in MyData
for (i in seq(1, 15)){
    MyData = MyData %>%
    dplyr::mutate(Drug =  ifelse(Plate == i, Drugs[i], Drug))
}

# Removing the Plate column
MyData = subset(MyData, select = -c(Plate)) 
```

Getting rid of the units for the doses and associating 0.5% to the null dose (it corresponds DMSO).
We can do that because a drug has either been measured in $n g$ or $\mu g$.

```{r}
MyData = MyData %>%
  dplyr::mutate(Dose =  ifelse(Dose %in% c("1ug", "1ng"), "1" , Dose))
MyData = MyData %>%
  dplyr::mutate(Dose =  ifelse(Dose %in% c("10ug","10ng"), "10" , Dose))
MyData = MyData %>%
  dplyr::mutate(Dose =  ifelse(Dose %in% c("100ug", "100ng"), "100" , Dose))
MyData = MyData %>%
  dplyr::mutate(Dose =  ifelse(Dose %in% c("1000ug", "1000ng"), "1000" , Dose))
MyData = MyData %>%
  dplyr::mutate(Dose =  ifelse(Dose %in% c("0.50%", "0.5%"), "0", Dose))
```

Renaming the columns and two of the stimulation for convenience.

```{r}
MyData = MyData %>% 
            dplyr::rename("dose" = "Dose",
                          "drug" = "Drug",
                          "stimulation" = "Stims") %>%
            dplyr::mutate(stimulation = gsub("GM-CSF", "GMCSF", stimulation),
                          stimulation = gsub("Unstim", "unstim", stimulation))
```

```{r}
head(MyData)
```

## Preprocessing of the features

Here use the asinh transform on the cytometry data and build a feature that is the difference between stimulated and un-stimulated samples .

```{r}
final_data = c()
metadata = setdiff(colnames(MyData), c("dose", "stimulation", "median"))

for(drug.name in unique(MyData$drug)){
  # Filtering the data with a specific drug
  MyData_drug = MyData %>% dplyr::filter(drug == drug.name) %>%
                  dplyr::mutate(conditions=paste0(stimulation, "_", dose, sep=""))
  # Getting the stims for this drug
  conditions = unique(MyData_drug$conditions)
  # asinh transformation
  MyData_drug = MyData_drug %>% dplyr::mutate(feature = asinh(median/5))
  # Removing the median column and calculating Stimulated - Unstimulated_0
  MyData_drug = MyData_drug[, names(MyData_drug) != "median"] %>%
      dplyr::select(-c(stimulation, dose)) %>%
      pivot_wider(names_from = conditions, values_from = feature)
  
  MyData_drug["unstim_-1"] = MyData_drug["unstim_0"] # Saving the asinh transform of unstim_0
  
  for (condition in conditions){
      MyData_drug[condition] = MyData_drug[condition] - MyData_drug["unstim_0"]
  }
  
  MyData_fin = MyData_drug %>% 
                  pivot_longer(-all_of(metadata), names_to="conditions",values_to="feature") %>%
                  dplyr::mutate(dose=as.double(sub(".*_(.*)", "\\1", conditions)),
                                stimulation=sub("_.*", "", conditions)) %>%
                  dplyr::select(-conditions)

  # Adding the new calculated features to the final dataset
  final_data = rbind(final_data, MyData_fin)
}
```

Saving the final preprocessed dataset.

```{r}
write.csv(final_data, paste(out_path, "preprocessed.csv", sep = "/"))
```

## Penalized Data

This is function taking a penalization matrix and the name of the stimulation associated with the matrix and returning the list of the features to be removed from the analysis.

```{r}
penalization_preprocessing <- function(penalization_matrix, stim.name){
  penalization_matrix <- penalization_matrix %>%
  `colnames<-`(c("population", colnames(penalization_matrix)[-1])) %>%
  dplyr::mutate(population = as.factor(population)) %>%
  gather("protein", "penalization", -population) %>% 
  dplyr::mutate(feature = paste(population, "_", protein, "_", stim.name, sep="")) %>% 
  dplyr::filter(penalization == 0) %>%
  dplyr::mutate(feature = sub("_[^_]*_", "_", feature))
}
```

Getting all the features to remove due to penalization on each stimulation.

```{r}
stims = c("IFNa","LPS","IL246","GMCSF")
discard = c()

for (stim in stims){
  new_penalization = penalization_preprocessing(read_xlsx(path_to_penalization_file, sheet = stim)[1:12], stim)
  discard <- c(discard, new_penalization$feature)
}
```

Removing penalized features from the unpenalized dataset.

```{r}
pen_final_data <- final_data %>%
                    dplyr::mutate(feature.name = paste(population, "_", reagent, "_", stimulation, sep="")) %>%
                    dplyr::filter(!feature.name %in% discard) %>%
                    dplyr::select(-feature.name)
```

Saving penalized dataset.

```{r}
write.csv(pen_final_data, paste(out_path, "preprocessed_pen.csv", sep = "/"))
```

